---
title: "pm566_wk7_lab"
author: "Yiping Li" 
output: github_document
date: "`r Sys.Date()`" 
always_allow_html: true
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(tidyverse)
library(httr)
library(stringr)
```

Question 1: How many sars-cov-2 papers?
Build an automatic counter of sars-cov-2 papers using PubMed. You will need to apply XPath as we did during the lecture to extract the number of results returned by PubMed in the following web address: https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2
```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,+]")
```

```{r}
stringr::str_extract(counts, "[0-9,]+")
```
Question 2: Academic publications on COVID19 and Hawaii
You need to query the following The parameters passed to the query are documented here.

Use the function httr::GET() to make the following query:

Baseline URL: https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi

Query parameters:

db: pubmed
term: covid19 hawaii
retmax: 1000
```{r}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(db='pubmed',
               term = 'covid19 hawaii',
               retmax= 1000)
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)
```
Question 3: Get details about the articles
The Ids are wrapped around text in the following way: <Id>... id number ...</Id>. we can use a regular expression that extract that information. Fill out the following lines of code:
```{r}
# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]</Id")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")
```
With the ids in hand, we can now try to get the abstracts of the papers. As before, we will need to coerce the contents (results) to a list using:

Baseline url: https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi

Query parameters:

db: pubmed
id: A character with all the ids separated by comma, e.g., “1232131,546464,13131”
retmax: 1000
rettype: abstract
Pro-tip: If you want GET() to take some element literal, wrap it around I() (as you would do in a formula in R). For example, the text "123,456" is replaced with "123%2C456". If you don’t want that behavior, you would need to do the following I("123,456").
```{r}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db='pubmed',
    id=paste(ids,collapse =','),
    retmax=1000,
    rettype='abstract'
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```

